### 2021-01-24 paper/arm.tex
FOrward reference in 2.1
Related work paragraph

We discuss \armeight{} using \emph{external global completion} (\EGC{})
\cite{alglave-git-alternate,armed-cats}, which is very close to our model;
see also \cite[\textsection B2.3.6]{arm-reference-manual}.

@@

\begin{example}
  The following execution is allowed by \armeight{}, but disallowed by
  \refdef{def:pomsets-ra}.  The coherence order between the writes can be
  witnessed by a separate thread, which we have elided.
  \begin{gather*}
    \PW{x}{2}\SEMI 
    \PR[\mRA]{x}{r}\SEMI
    \PW{y}{1} \PAR
    \PW{y}{2}\SEMI
    \PW[\mRA]{x}{1}
    \\
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
        \rf{a}{b}
        \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}
  \end{gather*}
  Under \EGC{}, this is explained by dropping the order
  $\DRP[\mRA]{x}{2}\xsync\DWP{y}{1}$, because $\DRP[\mRA]{x}{2}$ is fulfilled
  by a relaxed write in the same thread.
  \begin{gather*}
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
        \rf{a}{b}
        % \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}  
  \end{gather*}

@@

To model such \emph{downgraded reads}, we use the uninterpreted symbols
$\Dx{\aLoc}$.

@@

@@

\begin{definition}[$\xRRD$]
  \label{def:pomsets-rr}
  Update \refdef{def:pomsets-trans} to:
  \begin{enumerate}
  \item[\ref{T3})]
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
    \\
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\FALSE/\RW]$ otherwise.
  \end{enumerate}
  \begin{enumerate}
  \item[\ref{L5})]
    %$\aTr{\cEvs}{\bForm}$ implies $\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm$,
    $\aTr{\cEvs}{\bForm}$ implies $\RW\limplies\bForm$,
  \end{enumerate}  
\end{definition}

@@

\refdef{def:pomsets-arm} combines all of the features of
\textsection\ref{sec:q}--\ref{sec:arm}.  For completeness, we repeat
\ref{L4}, which is unchanged from \refdef{def:pomsets-trans}.
%Only \ref{L5} and \ref{T3} are changed from \refdef{def:pomsets-down}.

\begin{definition}[$\xCO$/$\xRASC$/$\xDGR$/$\xRRD$]
  \label{def:pomsets-arm}
  Update \refdef{def:pomsets-trans} to:
  \begin{enumerate}
  \item[\ref{S3})]
    $\labelingForm(\aEv)$ implies $\QS{\aLoc}{\amode}\land\aExp{=}\aVal$,
  \item[\ref{L3})]
    $\labelingForm(\aEv)$ implies $\QL{\aLoc}{\amode}$.
  \item[\ref{T3})]
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
    \\
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\FALSE/\RW]$ otherwise.
  \end{enumerate}
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies $(\Qw{\aLoc}\limplies\aExp{\neq}\aVal)\land \DS{\aLoc}{\amode}\bForm\noSUB{[\aExp/\aLoc]}$,
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qw{\aLoc}\land \DS{\aLoc}{\amode}\bForm\noSUB{[\aExp/\aLoc]}$.
  \item[\ref{L4})]
    $\aTr{\bEvs}{\bForm}$ implies $\aVal{=}\aReg\limplies\bForm$, 
  \item[\ref{L5})]
  %   $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land(\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm)$,
  % \item[\ref{L6})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode}
    \land (\RW\limplies\bForm)$.
  \end{enumerate}  
\end{definition}
\begin{proposition}
  Every \armeight{} execution is allowed by \refdef{def:pomsets-arm}.
  \begin{proof}[Proof Sketch]
    By \EGC{}, an \armeight{} execution linearization of local-order and
    per-location program order.  Every such linearization is also a valid
    pomset under \refdef{def:pomsets-arm}.
  \end{proof}
\end{proposition}

>>>>

We discuss \armeight{} using \emph{external global completion} (\EGC{})
\cite{alglave-git-alternate} %,armed-cats},
\cite[\textsection B2.3.6]{arm-reference-manual}
which is very close to our model.

@@

\begin{example}
  \label{ex:dgr1}
  The following execution is allowed by \armeight{}, but disallowed by
  \refdef{def:pomsets-ra}.  The coherence order between the writes can be
  witnessed by a separate thread, which we have elided.
  \begin{gather*}
    \PW{x}{2}\SEMI 
    \PR[\mRA]{x}{r}\SEMI
    \PW{y}{1} \PAR
    \PW{y}{2}\SEMI
    \PW[\mRA]{x}{1}
    \\
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
        \rf{a}{b}
        \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}
  \end{gather*}
  Under \EGC{}, this is explained by dropping the order
  $\DRP[\mRA]{x}{2}\xsync\DWP{y}{1}$, because $\DRP[\mRA]{x}{2}$ is fulfilled
  by a relaxed write in the same thread.
  \begin{gather*}
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
        \rf{a}{b}
        % \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}  
  \end{gather*}

@@

To model such \emph{downgraded reads}, we use the uninterpreted symbols
$\Dx{\aLoc}$.  Load actions that requiring downgrading introduce $\Dx{\aLoc}$.
Relaxed stores on $\aLoc$ substitute true for $\Dx{\aLoc}$, whereas
synchronizing stores substitute false for $\Dx{\aLoc}$.

@@

\begin{example}
  Revisiting \refex{ex:dgr1}.
  \begin{gather*}
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
        \rf{a}{b}
        % \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}  
  \end{gather*}  
\end{example}

@@

\begin{definition}[$\xRRD$]
  \label{def:pomsets-rr}
  Update \refdef{def:pomsets-trans} to:
  \begin{enumerate}
  \item[\ref{T3})]
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
    \\
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\FALSE/\RW]$ otherwise.
  \end{enumerate}
  \begin{enumerate}
  \item[\ref{L5})]
    % $\aTr{\cEvs}{\bForm}$ implies $\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm$,
    $\aTr{\cEvs}{\bForm}$ implies $\RW\limplies\bForm$,
  \end{enumerate}  
\end{definition}

@@

\refdef{def:pomsets-arm} combines all of the features of
\textsection\ref{sec:q}--\ref{sec:arm}.  For completeness, we repeat
\ref{L4}, which is unchanged from \refdef{def:pomsets-trans}.
% Only \ref{L5} and \ref{T3} are changed from \refdef{def:pomsets-down}.

\begin{definition}[$\xCO$/$\xRASC$/$\xDGR$/$\xRRD$]
  \label{def:pomsets-arm}
  Update \refdef{def:pomsets-trans} to:
  \begin{enumerate}
  \item[\ref{S3})]
    $\labelingForm(\aEv)$ implies $\QS{\aLoc}{\amode}\land\aExp{=}\aVal$,
  \item[\ref{L3})]
    $\labelingForm(\aEv)$ implies $\QL{\aLoc}{\amode}$.
  \item[\ref{T3})]
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
    \\
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\FALSE/\RW]$ otherwise.
  \end{enumerate}
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies $(\Qw{\aLoc}\limplies\aExp{\neq}\aVal)\land \DS{\aLoc}{\amode}\bForm\noSUB{[\aExp/\aLoc]}$,
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qw{\aLoc}\land \DS{\aLoc}{\amode}\bForm\noSUB{[\aExp/\aLoc]}$.
  \item[\ref{L4})]
    $\aTr{\bEvs}{\bForm}$ implies $\aVal{=}\aReg\limplies\bForm$, 
  \item[\ref{L5})]
    % $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land(\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm)$,
    % \item[\ref{L6})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode}
    \land (\RW\limplies\bForm)$.
  \end{enumerate}  
\end{definition}

Every \armeight{} execution is allowed by \refdef{def:pomsets-arm}.  The
proof of this fact is simplified by the recent characterization of
\armeight{} in terms of \EGC{} \cite[\textsection
B2.3.6]{arm-reference-manual}.  Under \EGC{}, an \armeight{} execution is a
linearization of per-location program order and a subset of local-order.
Every such linearization is also a valid pomset under
\refdef{def:pomsets-arm}.
